---
title: "MIS 306 Data Analysis | Competition"
author: "Group Members:<br>• Mert Özdemir<br>• Berdan Berke Baş<br>• Batuhan Karaköy<br>• Yunus Emre Işıktan"
subtitle: "NKE Stock Analysis - Group 3"
date: "2025-05-22"
output: slidy_presentation
---

```{r setup, include=FALSE, error=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(quantmod)
library(xts)
library(lubridate)
```

```{r include=FALSE}
load("djdata.RData")
```

```{r include=FALSE}
colnames(djdata)
colnames(djdata.df)
```

```{r include=FALSE}
symbols <- colnames(djdata)
symbols
symbols[-15]

set.seed(20022025)
sample(symbols[-15], 16)

set.seed(20022025)
lst <- sample(symbols[-15], 16)

lst[10]
djdata[,lst[10]]

getSymbols(lst[10])
```

---

# Objective
This project analyzes Nike Inc. (NKE) stock performance (2007–2025) to identify patterns, build forecasting models, and generate investment insights. Key components include:

<h4>1. Data Exploration</h4>

- Analyze daily/weekly/monthly returns for trends and seasonality

- Compare weekday distributions (Monday–Friday) using statistical tests

- Examine volatility patterns and trading volume changes

<h4>2. Model Development</h4>

- Test stationarity and apply differencing when needed

- Fit multiple ARIMA configurations (ARIMA(1,1,1), ARIMA(2,1,1))

- Implement automated model selection via auto.arima

- Generate forecasts using optimal models

<h4>3. Validation & Insights</h4>

- Evaluate model fit through residual diagnostics

- Compare performance using AIC/BIC and error metrics

- Identify reliable patterns for trading strategies

Goal: Provide actionable intelligence about NKE's price behavior and robust forecasting tools for investors.

---

# Visual Summaries
## Plot for opening, closing prices, high and low values and the volume
```{r echo=FALSE}
plot(as.zoo(NKE))
```

Nike’s stock exhibited strong growth from 2009 to 2021, peaking with high investor confidence. The post-2021 decline seen across all price indicators and the adjusted closing price suggests a broader market shift or internal performance issues. Trading volume spikes throughout the timeline point to periods of heightened interest, possibly tied to financial reports or economic events.

## Line Chart (2007 - 2025)
```{r echo=FALSE}
plot(NKE[,-5])
```

This chart shows the stock price of Nike (NKE) from January 2007 to May 2025. The price increased slowly until around 2013, then started to rise more quickly. It reached its highest point around the end of 2021, going above `$150`. After that, the price dropped a lot and continued to fall with some small increases in between. By 2025, the price is close to `$50`. Overall, the chart shows strong growth for many years, followed by a big decline after the peak.

---

```{r include=FALSE}
# First few days without labelling.
head(wday(index(NKE), label=F))

# First few days with labelling..
head(wday(index(NKE), label=T))
```

```{r include=FALSE}
# result is factor object..
str(head(wday(index(NKE), label=T)))

is.factor(head(wday(index(NKE), label=T)))

# Get the first few days as character (instead of factor)
paste(head(wday(index(NKE), label=T)))
```

```{r include=FALSE}
# days of year
yday(Sys.Date())

yday(index(first(NKE["200303"], "1 week")))

# days of quarter
qday(Sys.Date())

head(qday(index(NKE)))

quarter(Sys.Date()) # quarter of the date

# days of month
head(mday(index(NKE)))

# month of first few obs.
head(month(index(NKE)))

# Convert to monthly data using to.monthly() function. Then get the first few months
head(month(index(to.monthly(NKE)), label=T))

# Just names only..
paste(head(month(index(to.monthly(NKE)), label=T)))
```

```{r include=FALSE}
# last weekdays of months and of course weekdaysdate
last_weekday_o_month <- function(dx) {
  ave( 
  paste(wday(dx, label=T, abbr=F)), 
  paste(month(dx, label=T, abbr = F)), 
  FUN = function(x) tail(x[ !(x %in% c("Saturday","Sunday")) ], 1) 
  ) 
}

last_weekdaydate_o_month <- function(dx){
  ave( 
    dx, 
# paste(month(dx, label=T, abbr = F)), 
    month(dx, label=T),
    year(dx),
    FUN = function(x) tail(x[ !(wday(x, label=T, abbr=F) %in% c("Saturday","Sunday")) ], 1) 
  )
}  

# Get the last 25 end of months weekdays Adjusted close value..
Ad(NKE[tail(unique(last_weekdaydate_o_month(index(NKE))), 15),])
```

```{r include=FALSE}
NKE_AD <- Ad(NKE)

head(NKE_AD[wday(index(NKE_AD))==2])

wday(head(NKE_AD[wday(index(NKE_AD))==2]), label=T)

wday(head(NKE_AD[wday(index(NKE_AD))==6]), label=T)

NKE_AD.ret <- diff(log(NKE_AD))

head(NKE_AD.ret[wday(index(NKE_AD.ret))==2])

head(merge(NKE_AD, NKE_AD.ret))

head(merge(NKE_AD, NKE_AD.ret)[wday(merge(NKE_AD, NKE_AD.ret))==2])
```

# Visual Summaries 
## Histogram: NKE returns (All Days | Monday | Friday)
```{r echo=FALSE, message=FALSE}
Sys.setlocale("LC_TIME", "C")
library(zoo)

par(mfrow = c(1, 3))

hist(NKE_AD.ret, breaks = 100, main = "All Days", xlab = "Returns")

hist(NKE_AD.ret[weekdays(index(NKE_AD.ret)) == "Monday"], breaks = 50, main = "Monday", xlab = "Returns")

hist(NKE_AD.ret[weekdays(index(NKE_AD.ret)) == "Friday"], breaks = 50, main = "Friday", xlab = "Returns")

par(mfrow = c(1, 1))
```

This figure shows the distribution of Nike’s daily returns, including all days, Mondays, and Fridays. All three histograms have a similar shape—most returns are close to zero, forming a sharp peak in the center. This means small daily changes are more common. Extreme values (both gains and losses) are rare but exist. Compared to Mondays, Friday returns seem slightly more spread out, which may suggest more market activity or news impact before the weekend. Overall, the returns appear to follow a normal-like distribution with slight differences by weekday.

---

## Empirical Cumulative Distribution of Fridays Return
```{r echo=FALSE}
plot(ecdf(as.vector(NKE_AD.ret[wday(index(NKE_AD.ret))==6])), main="Friday ECDF")
```

Most of the returns are between `-0.05` and `0.05`, with the curve rising steeply in that range. This means that small returns are the most common. The flat areas on both sides show rare extreme values, either losses or gains. Overall, the returns are centered around zero, and the shape of the curve suggests a normal-like distribution.

## Comparison of Empirical Cumulative Distributions of Monday (red) and Fridays (black) Return
```{r echo=FALSE}
plot(ecdf(as.vector(NKE_AD.ret[wday(index(NKE_AD.ret))==6])), main="ECDF Distributions")
lines(ecdf(as.vector(NKE_AD.ret[wday(index(NKE_AD.ret))==2])), col="red")
```

This ECDF plot compares the return distributions of two groups, likely all days (black) and Fridays (red). Both lines are very close to each other, meaning their return patterns are similar. Most returns are between `-0.05` and `0.05`, and the sharp rise around zero shows that small returns are the most common. The few extreme values on both sides show rare large losses or gains. Overall, the distributions are almost the same, with only small differences.

---

# Mean Returns
## Mean of the weekday returns
This table shows the average daily returns of Nike stock for each weekday. Monday has a small negative mean return (`-0.000448`), while all other days have positive mean returns. Thursday has the highest average return (`0.000759`), followed by Tuesday. This suggests that, on average, Nike performs slightly worse on Mondays and slightly better on Thursdays. Overall, the differences are small, but they show a possible weekday effect in returns.

```{r echo=FALSE}
mean.ret <- cbind(
mean(NKE_AD.ret[wday(index(NKE_AD.ret))==2], na.rm=T),
mean(NKE_AD.ret[wday(index(NKE_AD.ret))==3], na.rm=T),
mean(NKE_AD.ret[wday(index(NKE_AD.ret))==4], na.rm=T),
mean(NKE_AD.ret[wday(index(NKE_AD.ret))==5], na.rm=T),
mean(NKE_AD.ret[wday(index(NKE_AD.ret))==6], na.rm=T))
colnames(mean.ret) <- levels(wday(index(NKE_AD.ret), label=T))[2:6]
rownames(mean.ret) <- "Mean Returns"

round(mean.ret, 6)
```
```{r echo=FALSE}
library(ggplot2)

mean_ret_df <- data.frame(
  Weekday = factor(
    colnames(mean.ret),
    levels = c("Mon", "Tue", "Wed", "Thu", "Fri")
  ),
  Mean_Return = as.numeric(mean.ret)
)

ggplot(mean_ret_df, aes(x = Weekday, y = Mean_Return, fill = Weekday)) +
  geom_bar(stat = "identity") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  labs(
    title = "Mean Stock Returns by Weekday (Mon-Fri)",
    x = "Weekday",
    y = "Mean Return"
  ) +
  scale_y_continuous(labels = scales::comma) +
  theme_minimal() +
  theme(legend.position = "none")
```

# NKE Monday Prices Line Chart
```{r echo=FALSE}
plot(NKE_AD[wday(index(NKE_AD))==2], main="NKE Mondays Prices", ylab="$")
```

This chart shows Nike’s stock prices on Mondays from January 2007 to May 2025. The prices rose steadily until around late 2021, reaching a peak above `$150`. After that, the trend reversed, and the prices started to fall. By 2025, the price dropped to nearly $60. This shows a long period of growth followed by a strong decline, with Monday prices reflecting the overall market behavior.

---

# NKE Monday Prices Zoo
```{r echo=FALSE}
# zoo type..
#plot(as.zoo(NKE[wday(index(NKE))==2]), main="NKE Mondays Prices", ylab="$", xlab="Date")

nke_mon <- NKE[wday(index(NKE)) == 2]

plot(as.zoo(nke_mon), main = "NKE Mondays Prices", ylab = "$", xlab = "Date", col = 1:ncol(nke_mon))

```

This figure shows six different panels of Nike’s Monday stock data from 2007 to 2025. Each panel represents a different variable:

-Top-left (black line): Adjusted closing prices. It shows a strong upward trend until late 2021, then a clear decline.

-Middle-left (red line): Opening prices. These follow a very similar pattern to the adjusted prices, peaking in 2021.

-Bottom-left (green line): Closing prices. The movement is also very close to the adjusted and opening prices.

-Top-right (blue line): High prices. This line represents the highest price on each Monday and follows the same rising and falling pattern.

-Middle-right (cyan line): Volume. It shows the number of shares traded. There are many spikes, especially after 2020, which may indicate market reactions to news or events.

-Bottom-right (purple line): Low prices. It shows the lowest value of the stock on each Monday, also peaking in 2021 and dropping afterwards.

Overall, all price-related charts show growth until around 2021, then a decline. The volume chart shows trading activity became more volatile in recent years. Mondays clearly reflect the general stock trend of Nike.

# Volume Log Transformation
```{r echo=FALSE}
nke_mon_vol_log <- log10(as.numeric(nke_mon[, "NKE.Volume"]))

plot(index(nke_mon), nke_mon_vol_log, type = "l", col = "steelblue",
     main = "Log-Transformed Volume (Mondays)", ylab = "log10(Volume)", xlab = "Date")
```

This chart shows the log-transformed trading volume of Nike stock on Mondays from 2007 to 2025. The y-axis uses a log scale to better show volume changes. In the early years, volume was higher and more volatile. Between 2015 and 2020, it stayed relatively low and stable. After 2022, volume started to increase again, with more sharp spikes near 2025. This suggests rising market activity and interest in recent years.

---

```{r include=FALSE}
nke_adj <- Ad(NKE)

nke_monthly_adj <- to.monthly(nke_adj, indexAt = "lastof", OHLC = FALSE)

nke_monthly_ret <- diff(log(nke_monthly_adj))

round(mean(nke_monthly_ret, na.rm = TRUE), 6)

```

```{r include=FALSE}
round((1+mean(nke_monthly_ret, na.rm=T))^12-1, 6)
```

```{r include=FALSE}
nke_yearly_adj <- to.yearly(nke_adj, indexAt = "lastof", OHLC = FALSE)

nke_yearly_ret <- diff(log(nke_yearly_adj))

round(mean(nke_yearly_ret, na.rm = TRUE), 6)

```

```{r include=FALSE}
head(nke_monthly_ret[month(index(nke_monthly_ret))==1], na.rm=T)
```

```{r include=FALSE}
mean(nke_monthly_ret[month(index(nke_monthly_ret))==1], na.rm=T)

mean(nke_monthly_ret[month(index(nke_monthly_ret))==2], na.rm=T)
```

# NKE Monthly Returns NKEr Plot (2017 - 2025)
```{r echo=FALSE}
barplot(nke_monthly_ret[month(index(nke_monthly_ret))==1])
```

This bar chart shows Nike’s January returns from 2007 to 2024. Most bars are below zero, meaning that January often had negative returns. The largest drop happened around 2009, likely during the financial crisis. Positive returns appear in some years like 2011, 2012, and 2023. Overall, January returns are mixed, but negative values seem more common, showing that the stock often performs weakly at the start of the year.

---

# Statistical Tests
## Kolmogorov-Smirnov Test
This Kolmogorov-Smirnov test compares the return distributions of Mondays and Tuesdays. The test statistic is `D = 0.067225`, and the p-value is `0.03324`. Since the p-value is below 0.05, the result is statistically significant. This means there is evidence that the return distributions for Mondays and Tuesdays are different, suggesting that returns behave differently on these two weekdays.

```{r warning=FALSE, echo=FALSE}
# Compare Mondays and Tuesdays..
ks.test(NKE_AD.ret[wday(index(NKE_AD.ret))==2], NKE_AD.ret[wday(index(NKE_AD.ret))==3])
```
This Kolmogorov-Smirnov test compares the return distributions of Mondays and Fridays. The test statistic is `D = 0.046412`, and the p-value is `0.2894`, which is quite high. This means there is no significant evidence to reject the null hypothesis. In other words, the return distributions for Monday and Friday are statistically similar, with no clear difference between them.

```{r warning=FALSE, echo=FALSE}
# Compare Mondays and Fridays..
ks.test(NKE_AD.ret[wday(index(NKE_AD.ret))==2], NKE_AD.ret[wday(index(NKE_AD.ret))==6])
```

---

## Time Series Decomposition
```{r echo=FALSE}
NKE.monthly.adj <- to.monthly(nke_adj, indexAt = "lastof", OHLC = FALSE)

NKE.ts <- ts(as.numeric(NKE.monthly.adj), start = c(2007, 1), frequency = 12)

NKE.stl <- stl(NKE.ts, s.window = "periodic")

plot(NKE.stl)

```

This plot shows the decomposition of Nike’s monthly adjusted prices into four parts: data, seasonal, trend, and remainder. The data panel shows the original price movement, with growth until 2021, followed by a decline. The seasonal panel reveals repeating yearly patterns, showing that seasonality exists in the data. The trend panel shows long-term growth peaking around 2021, then dropping. The remainder shows random noise or irregular movements. Overall, the stock price is influenced by both seasonality and long-term trends.

---

## Forecasting
Since we both have trend and seasonality in our data we will use Holt Winter's Forecast Method.

```{r echo=FALSE}
library(forecast)
library(ggplot2)

NKE.hw <- ets(NKE.ts, model = "AAA")

NKE.hw.forecast <- forecast(NKE.hw, h = 12)

autoplot(NKE.hw.forecast) +
  ggtitle("NKE Monthly Adjusted Price Forecast (Holt-Winters Method)")
```

This chart shows the forecast of Nike’s monthly adjusted prices using the Holt-Winters method. The black line represents historical prices, which rose sharply until 2021 and then declined. The blue line shows the forecast for the next 12 months, while the shaded area indicates the confidence intervals. The forecast suggests a slight upward trend, but the wide shaded area shows there is uncertainty. Overall, the model expects possible recovery, but with some risk.

```{r echo=FALSE}
summary(NKE.hw)
```

This summary shows the results of the Holt-Winters (ETS AAA) model applied to Nike’s monthly prices. The model uses additive error, trend, and seasonality. The smoothing parameters (`alpha = 0.9971`) suggest the model gives strong weight to recent values. The initial seasonal states show clear seasonal effects. The AIC (`1983.67`) and RMSE (`5.54`) indicate the model’s fit. Error measures like MAPE (`6.94%`) show the average forecast error is relatively low. Overall, the model fits the data well and captures both trend and seasonal patterns.

---

# Smoothing
## Adjusted Close
```{r warnings = FALSE, message=FALSE, echo=FALSE}
library(zoo)
library(fpp3)
library(mgcv)

autoplot(NKE_AD) +
  geom_smooth(method = "loess", span = 0.3, color = "red") +
  labs(
    title = "Loess Smoothing: Adjusted Close Prices",
    x = "Date",
    y = "Adjusted Close ($)"
  ) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(linewidth = 1, colour = "grey80"))
```

This chart shows Nike’s adjusted close prices over time, with daily price movements in black and a red Loess smoothing line to highlight the trend. The Loess technique helps reduce short-term fluctuations and reveals the general pattern more clearly. From 2007 to around 2021, the red line shows a strong upward trend, peaking sharply around 2020. After that, the trend turns downward, indicating a long-term decline. Despite ongoing variations in daily prices, the smoothed line makes it easier to see the overall rise and fall in stock performance.

---

## High Prices
```{r warnings = FALSE, echo=FALSE, message = FALSE}
autoplot(NKE$NKE.High) +
  geom_smooth(method = "loess", span = 0.3, color = "red") +
  labs(
    title = "Loess Smoothing: High Prices",
    x = "Date",
    y = "High Price ($)"
  ) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(linewidth = 1, colour = "grey80"))
```

```{r echo=FALSE}
autoplot(NKE$NKE.High) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3), color = "red") +
  labs(
    title = "Polynomial Regression Smoothing: High Prices",
    x = "Date",
    y = "High Price ($)"
  ) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(linewidth = 1, colour = "grey80"))
```

These charts show Nike’s high stock prices over time, using two different smoothing methods.

- The first chart uses Loess smoothing. The black line shows the actual high prices, while the red line reveals a smooth trend. It clearly shows a steady increase until 2021, followed by a noticeable decline. This method captures short-term movements while showing the overall trend.

- The second chart uses polynomial regression smoothing. The red line is smoother and more general. It shows a slow rise from 2010 to 2021, then a decline, but it doesn't follow the short-term fluctuations as closely.

Overall, both methods show a long-term rise and fall in prices, but Loess gives a more detailed view, while polynomial regression gives a broader trend. 

---

## Low Prices
```{r warnings = FALSE, message=FALSE, echo=FALSE}
library(patchwork)

plot1 <- autoplot(NKE$NKE.Low) +
  geom_smooth(method = "loess", span = 0.1, color = "red") +
  labs(
    title = "Loess Smoothing (span=0.1): Low Prices",
    x = "Date",
    y = "Low Price ($)"
  ) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(linewidth = 1, colour = "grey80"))

plot2 <- autoplot(NKE$NKE.Low) +
  geom_smooth(method = "loess", span = 0.3, color = "red") +
  labs(
    title = "Loess Smoothing (span=0.3): Low Prices",
    x = "Date",
    y = "Low Price ($)"
  ) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(linewidth = 1, colour = "grey80"))

plot3 <- autoplot(NKE$NKE.Low) +
  geom_smooth(method = "loess", span = 0.5, color = "red") +
  labs(
    title = "Loess Smoothing (span=0.5): Low Prices",
    x = "Date",
    y = "Low Price ($)"
  ) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(linewidth = 1, colour = "grey80"))

plot4 <- autoplot(NKE$NKE.Low) +
  geom_smooth(method = "loess", span = 0.7, color = "red") +
  labs(
    title = "Loess Smoothing (span=0.7): Low Prices",
    x = "Date",
    y = "Low Price ($)"
  ) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(linewidth = 1, colour = "grey80"))

(plot1 + plot2) / (plot3 + plot4)
```

This chart compares Loess smoothing with different span values applied to Nike’s low prices. The black lines show actual prices, and red lines show smoothed trends. With span = 0.1, the red line closely follows the data, showing more detail. As the span increases to 0.3, 0.5, and 0.7, the red line becomes smoother and less sensitive to short-term changes. Larger spans highlight the overall trend better, while smaller spans capture more local fluctuations. Overall, the choice of span affects how much of the short-term movement is shown in the trend.

---

# ARMA & ARIMA
## Stationarity Testing and Autocorrelation Analysis
The ADF test shows Nike's stock prices are not stable over time (Dickey-Fuller = -0.43, p-value = 0.99). This means the prices follow trends that change. Since the test says we need to difference the data once (ndiffs() = 1), we should calculate daily price changes. After this change, we can check again if the data becomes stable. This step is important before using ARIMA models, which need stable data to work properly. The very high p-value proves the prices need this adjustment.

```{r warning=FALSE, echo=FALSE}
library(tseries)
nke_close <- log(Cl(NKE))
plot(nke_close, main = "Log Transformed NKE Prices")
#why did we take the logarith of Closing prices?
#-Reduces heteroskedasticity (volatility clustering).
#-Makes multiplicative trends additive.
adf.test(as.numeric(nke_close))
ndiffs(nke_close)#since this is return us '1' we we differenciate only 1 times

acf(nke_close)  # Plot autocorrelation
pacf(nke_close) # Partial autocorrelation
```

The log-transformed NKE prices graph shows the stock's price movement from 2007 to 2025. The upward trend until 2021 indicates strong growth, followed by a recent decline. This long-term pattern suggests the data needs differencing to remove trends before analysis.

The ACF (Autocorrelation Function) plot reveals how prices correlate with their past values. The gradual decline in correlation values across lags confirms the presence of trends, supporting the need for differencing to achieve stationarity.

The PACF (Partial Autocorrelation Function) plot identifies direct relationships between current prices and specific past prices. The significant spike only at lag 1 suggests an AR(1) component might be appropriate when modeling the differenced series. Together, these plots confirm the non-stationary nature of the data and guide appropriate model selection.

---

## ARIMA(1,1,1) Model Fitting and Residual Diagnostics
The ARIMA(1,1,1) model results show Nike's stock prices have a strong connection between consecutive days, with an AR(1) coefficient of 0.65 and an MA(1) coefficient of -0.69. The model fits the data well, as shown by the small error values (RMSE = 0.019) and the very low sigma squared value (0.00036). The ACF1 value close to zero (-0.0036) means the model has successfully captured all important patterns in the data, leaving only random noise in the residuals. With a high log likelihood (11,769) and low AIC (-23,532), this model provides reliable forecasts for Nike's stock price movements. The small prediction errors and clean residual plots confirm this is a good quality model for analyzing and forecasting Nike's stock performance.

```{r, echo=FALSE}
fit_arma111 <- arima(nke_close, order = c(1, 1, 1))
summary(fit_arma111)  # Changed from summary(fit)

plot(residuals(fit_arma111), type = "l", main = "Residuals of ARIMA(1,1,1) Model")  # Changed from residuals(fit)
acf(residuals(fit_arma111), main = "ACF of Residuals")  
pacf(residuals(fit_arma111), main = "PACF of Residuals")
```

These three charts help us understand the performance of the AR(1) model fitted to Nike’s return series.

- Residuals of ARIMA(1,1,1) Model chart shows the residuals over time. The values are centered around zero with small fluctuations, which is a good sign. However, there are a few spikes, showing some outliers or volatility in the data.

- The ACF plot shows the correlation between residuals at different time lags. All the bars (autocorrelations) fall within the blue confidence lines, meaning they are not significantly different from zero. This is exactly what we want to see - it tells us the model has successfully removed all predictable patterns from the data. The random spikes within the blue lines confirm the residuals now behave like pure random noise, with no remaining correlations that the model could have captured.

- The PACF plot examines direct relationships between residuals at specific lags. Nearly all points here fall very close to zero and stay within the confidence bands. While there are a couple of very small spikes just outside the bands (around lags 5 and 15), these are minor and likely due to random chance rather than any real pattern. Like the ACF, this confirms our model has done a good job - there are no important correlations left in the residuals that we failed to account for in our ARIMA model.

---

## ARIMA(1,1,2) Model Estimation and Diagnostic Checking
The ARIMA(1,1,2) model shows good results for Nike's stock prices. The model coefficients (AR1 = 0.65, MA1 = -0.70, MA2 = 0.005) suggest the current price is influenced by recent prices and past shocks. The very small sigma² value (0.00036) and high log likelihood (11769) indicate a strong fit. Error measures are excellent, with low RMSE (0.019) and MAE (0.0126), meaning the model's predictions are close to actual prices. The residual mean is near zero (0.0004), and the Box-Ljung test (p-value = 0.136) confirms the residuals contain no remaining patterns - they're just random noise. The ACF and PACF plots (not shown here) would likely confirm this clean residual behavior. Overall, this model performs slightly better than the simpler ARIMA(1,1,1) version, though the small MA2 coefficient suggests the second moving average term may not be adding much value.

```{r, echo=FALSE}
fit_arma112 <- arima(nke_close, order = c(1, 1, 2))
summary(fit_arma112)

acf(residuals(fit_arma112), main = "ACF of Residuals")  
pacf(residuals(fit_arma112), main = "PACF of Residuals")

mean(residuals(fit_arma112))
Box.test(residuals(fit_arma112), type = "Ljung-Box", lag = 10)
```

- The ACF plot shows the correlation between residuals at different time intervals. All the bars (correlation values) stay completely within the blue confidence lines, which means they are not significantly different from zero. This is exactly what we want to see - it tells us the model has successfully removed all important patterns from the data. The residuals now behave like random noise, with no remaining correlations that the model could have captured. This clean ACF plot gives us confidence that our ARIMA model is working properly.

- The PACF plot examines the direct relationship between residuals at specific time lags. Nearly all points here fall very close to zero and stay well within the confidence bands. While there are two very small spikes around lags 5 and 15 that slightly cross the blue lines, these are minor and likely due to random chance rather than any real pattern. Like the ACF, this confirms our model has done a good job - there are no significant correlations left in the residuals that we failed to account for in our ARIMA model. The tiny spikes we see are small enough to ignore for practical purposes.

---

## ARIMA(2,1,1) Model Estimation and Diagnostic Checking
The ARIMA(2,1,1) model provides a good fit for Nike's stock prices. The coefficients (AR1 = 0.67, AR2 = 0.007, MA1 = -0.72) show that current prices are mainly influenced by the most recent price (AR1) and recent shocks (MA1), while the second autoregressive term (AR2) has minimal impact. The model fits well, with a very small error variance (sigma² = 0.00036) and high log likelihood (11769). Prediction errors are low (RMSE = 0.019, MAE = 0.0126), indicating accurate forecasts. The Box-Ljung test (p-value = 0.133) confirms the residuals contain no remaining patterns, behaving like random noise. While this model performs similarly to the ARIMA(1,1,1), the near-zero AR2 coefficient suggests the second autoregressive term may not be necessary, making the simpler model potentially preferable.

```{r, echo=FALSE}
fit_arma211 <- arima(nke_close, order = c(2, 1, 1))
summary(fit_arma211)

acf(residuals(fit_arma211), main = "ACF of Residuals")  
pacf(residuals(fit_arma211), main = "PACF of Residuals")

Box.test(residuals(fit_arma211), type = "Ljung-Box", lag = 10)
```

- The ACF plot shows the autocorrelation of residuals at different lags. In this plot, all the bars after lag 0 are within the blue dashed lines, which represent the 95% confidence interval. This means there is no significant autocorrelation remaining in the residuals. Therefore, the model appears to have captured the time-related structure in the data well.

- The PACF plot shows the partial autocorrelation of residuals. Most of the bars are within the blue dashed lines, with only a few small spikes slightly outside. These small spikes are not strong enough to indicate a serious problem. This suggests that the residuals do not have a clear pattern left, and the model has likely captured the important information in the data.

---

## ARIMA(2,1,2) Model Implementation and Residual Diagnostics
The ARIMA(2,1,2) model results show some technical issues but decent overall performance. While the coefficients (AR1=0.055, AR2=0.354, MA1=-0.103, MA2=-0.379) suggest the model captures price patterns, the "NaN" standard errors indicate potential estimation problems - these values should normally be calculated. Despite this, the model shows good fit with small error variance (sigma²=0.00036) and high log likelihood (11769). Prediction errors remain low (RMSE=0.019, MAE=0.0126), similar to simpler models. The Box-Ljung test (p-value=0.1345) confirms the residuals contain no significant patterns, behaving like random noise. However, the estimation issues suggest this complex model may be overfitting the data, making the simpler ARIMA(1,1,1) potentially more reliable for forecasting Nike's stock prices.

```{r warning=FALSE, echo=FALSE}
fit_arma212 <- arima(nke_close, order = c(2, 1, 2))
summary(fit_arma212)

acf(residuals(fit_arma212), main = "ACF of Residuals")  
pacf(residuals(fit_arma212), main = "PACF of Residuals")

Box.test(residuals(fit_arma212), type = "Ljung-Box", lag = 10)
```

- The ACF plot shows all correlation values (vertical bars) staying within the blue confidence lines. This means there's no significant correlation between residuals at any time lag - they appear completely random. This clean result tells us the ARIMA model has successfully captured all predictable patterns in Nike's stock prices. The random, small bars within the confidence bands confirm the residuals now behave like pure noise, which is exactly what we want to see in a well-fitted time series model.

- The PACF plot shows nearly all points very close to zero, well within the confidence bands. While there are two tiny spikes around lags 5 and 15 that slightly cross the blue lines, these are so small they likely occurred by chance. Like the ACF, this confirms our model has done its job well - there are no important patterns left in the residuals. The minor spikes we see are insignificant and don't suggest any missing structure in our model. Together, these plots give us confidence that the ARIMA model is working properly.

---

## Optimal ARIMA Model Selection and Comparison
The auto.arima function selected an ARIMA(1,1,1) model as the best fit for Nike's stock prices. The model coefficients (AR1=0.6475, MA1=-0.6932) show that current prices are influenced by both the previous day's price and past random shocks. The small sigma² value (0.0003617) and high log likelihood (11769.13) indicate excellent model fit. Error measures are very low (RMSE=0.019, MAE=0.0126), meaning the model's predictions closely match actual prices. The ACF1 value near zero (-0.0036) confirms the residuals have no remaining patterns. With the lowest AIC score (-23532.25) among compared models, this ARIMA(1,1,1) is statistically the best choice for forecasting Nike's stock movements, balancing accuracy with simplicity.

```{r, echo=FALSE}
library(forecast)
auto_fit <- auto.arima(nke_close, seasonal = FALSE)
summary(auto_fit)

acf(residuals(auto_fit), main = "ACF of Auto ARIMA Residuals")
pacf(residuals(auto_fit), main = "PACF of Auto ARIMA Residuals")

# AIC comparison
AIC(auto_fit,fit_arma111 , fit_arma112, fit_arma212)
```

- The ACF plot shows all correlation values (vertical bars) falling completely within the blue confidence lines. This means there is no significant correlation between residuals at any time lag - they appear completely random. This clean pattern indicates our auto-selected ARIMA model has successfully removed all predictable patterns from the data. The residuals now behave like pure noise, which is exactly what we want to see in a well-fitted model.

- The PACF plot shows nearly all points very close to zero, well within the confidence bands. While there are two tiny spikes around lags 5 and 15 that slightly cross the blue lines, these are so small they likely occurred by chance. This confirms our model has effectively captured all important price patterns. The minor spikes we see don't represent any real missing structure and can be safely ignored for forecasting purposes.

- The table compares four different ARIMA models for Nike's stock prices. Both the auto-selected ARIMA(1,1,1) and manually specified ARIMA(1,1,1) show identical AIC scores (-23532.25), indicating equally good performance. The more complex models (ARIMA(1,1,2) and ARIMA(2,1,2)) have worse (higher) AIC values (-23530.35 and -23528.27 respectively), despite having more parameters. This tells us two important things: (1) the simpler ARIMA(1,1,1) model works just as well as the automatically selected one, and (2) adding extra parameters doesn't improve the model enough to justify the added complexity. Therefore, the ARIMA(1,1,1) is the best choice - it's simple but works very well for forecasting Nike's stock prices.

---

## Forecasting
```{r, echo=FALSE, message=FALSE, results='hide'}
forecast_values <- forecast(auto_fit, h = 300)
print(forecast_values)

plot(forecast_values, 
     main = "NKE Closing Price Forecast (Auto ARIMA)",
     xlab = "Time",
     ylab = "Prices")
```

The forecast graph shows Nike's future stock price predictions using the Auto ARIMA model. The black line represents historical prices, while the blue line shows the forecasted values. The shaded blue area indicates the prediction's confidence range - wider areas mean more uncertainty. The forecast suggests prices will remain relatively stable with small fluctuations, neither sharply rising nor falling. However, the confidence band gradually widens over time, showing predictions become less certain further into the future. This is normal in stock forecasting. The model expects Nike's prices to follow their recent pattern rather than making dramatic moves.

---

## Monthly NKE Price Analysis: ARIMA Model Selection and Forecasting
Comperison of two ARIMA models for Nike's monthly stock prices. The ARIMA(1,1,0) model shows a small negative autoregressive effect (AR1 = -0.0075) with drift (0.0072), suggesting prices tend to drift upward very slightly. The ARIMA(1,1,1) model has stronger coefficients (AR1 = 0.5184, MA1 = -0.5667) and a smaller drift (0.0035). Both models have similar error levels (sigma² ≈ 0.0056), but the ARIMA(1,1,0) performs slightly better with lower AIC (-511.19 vs -509.88) and BIC (-501.01 vs -496.31) values. This means the simpler ARIMA(1,1,0) model is likely better for forecasting, as it explains the data nearly as well with fewer parameters. The log likelihood values (258.6 vs 258.94) are very close, confirming both models fit the data reasonably well.

```{r warning=FALSE, echo=FALSE, results='hide'}
nke_monthly_close <- Cl(to.monthly(nke_close))

ggtsdisplay(nke_monthly_close, theme=theme_bw())

ggtsdisplay(diff(nke_monthly_close), theme=theme_bw())

auto_fit_m <-auto.arima(nke_monthly_close)

nke_monthly_close <- tsibble(nke_close=as.vector(coredata(nke_monthly_close)), 
                    date = yearmonth(zoo::index(nke_monthly_close)))

nke_monthly_close |> 
  gg_tsdisplay(nke_close, plot_type='partial')

nke_monthly_fit <- nke_monthly_close |> 
  model(arima110 = ARIMA(nke_close ~ 1+pdq(1,1,0) + PDQ(0,0,0)),
        arima011 = ARIMA(nke_close ~ 1+pdq(0,1,1)+ PDQ(0,0,0)),
        arima111 = ARIMA(nke_close ~ 1+pdq(1,1,1)+ PDQ(0,0,0)),
        stepwise = ARIMA(nke_close),
        search = ARIMA(nke_close, stepwise=FALSE))

nke_monthly_fit |> pivot_longer(everything(), 
                         names_to = "Model name",
                         values_to = "Orders")
report(nke_monthly_fit)

report(nke_monthly_fit |> dplyr::select(arima110))
report(nke_monthly_fit |> dplyr::select(arima111))

glance(nke_monthly_fit) |> 
  arrange(AICc) |> 
  dplyr::select(.model:BIC)

nke_monthly_fit |>
  dplyr::select(stepwise) |>
  gg_tsresiduals()

augment(nke_monthly_fit) |>
  filter(.model=='stepwise') |>
  features(.innov, ljung_box, lag = 10, dof = 1)

nke_monthly_fit |>
  forecast(h=16) |>
  filter(.model=='stepwise') |>
  autoplot(nke_monthly_close)

```

1. Original Time Series and ACF/PACF (Plot 1):
The first graph shows the original time series of Nike’s stock prices, which displays a clear upward trend over time. This means the data is non-stationary. The ACF plot shows a slow decrease in autocorrelation, which is a common sign of non-stationary time series. The PACF plot shows a strong spike at lag 1 and smaller spikes afterward. These results suggest that the time series has strong dependencies and needs to be differenced to become stationary.

2. Differenced Series and ACF/PACF (Plot 2):
The second graph shows the differenced series, likely the returns. This series fluctuates around zero and appears more stable. The ACF and PACF plots of the differenced series show no significant spikes beyond the confidence bands. This indicates that the differencing process worked well and the series is now stationary, which is important for time series modeling.

3. Fitted Model and ACF/PACF (Plot 3):
This plot shows the fitted values from the model and the corresponding ACF and PACF. The fitted series follows the trend of the original data well. The ACF and PACF again show high autocorrelation at the first lag and a gradual decrease, suggesting that the model captured the trend but some autocorrelation still exists. This can happen in models that do not fully remove all dependencies.

4. Residual Diagnostics (Plot 4):
This graph shows the residuals from the model, which appear to fluctuate randomly around zero. The ACF plot of the residuals shows that most autocorrelations are close to zero and within the confidence bands. The histogram of the residuals is roughly bell-shaped, indicating that they are approximately normally distributed. This is a good sign and suggests the model’s residuals behave like white noise.

5. Forecast Plot (Plot 5):
The last graph shows the forecast of Nike’s stock prices. The black line represents the historical data, while the blue shaded areas show the forecast with 80% and 95% confidence intervals. The forecast line is relatively flat, with some uncertainty as time goes forward. This means the model expects the stock price to stay stable in the near future, but with a possible range of outcomes.

---

## Weekly NKE Price Analysis: ARIMA Model Selection and Forecasting
Comperison of two ARIMA models for Nike's weekly stock prices. The ARIMA(1,1,0) model shows a small negative effect (-0.1159) from the previous week's price, with a tiny upward trend (0.0018). The ARIMA(1,1,1) model has a stronger negative effect (-0.5165) from past prices and accounts for random shocks (0.4026), with a slightly stronger trend (0.0025). Both models fit the data well, with nearly identical error levels (sigma² ≈ 0.0016). However, the ARIMA(1,1,1) performs slightly better, with a lower AIC (-3448.91 vs -3446.96) and higher log likelihood (1728.45 vs 1726.48), meaning it explains the data better despite being more complex. The small differences suggest either model could work, but the ARIMA(1,1,1) may capture more patterns in the weekly price changes.

```{r echo=FALSE, warning=FALSE, results='hide'}
nke_weekly <- to.weekly(NKE)
nke_weekly_close <- log(Cl(nke_weekly))

ggtsdisplay(nke_weekly_close, theme=theme_bw())

ggtsdisplay(diff(nke_weekly_close), theme=theme_bw())

auto_fit_m <-auto.arima(nke_weekly_close)

nke_weekly_close <- tsibble(nke_close=as.vector(coredata(nke_weekly_close)), 
                    date = yearweek(zoo::index(nke_weekly_close)))

nke_weekly_close |> 
  gg_tsdisplay(nke_close, plot_type='partial')

nke_weekly_fit <- nke_weekly_close |> 
  model(
        arima110 = ARIMA(nke_close ~ 1+pdq(1,1,0) + PDQ(0,0,0)),
        arima011 = ARIMA(nke_close ~ 1+pdq(0,1,1)+ PDQ(0,0,0)),
        arima111 = ARIMA(nke_close ~ 1+pdq(1,1,1)+ PDQ(0,0,0)),
        stepwise = ARIMA(nke_close),
        search = ARIMA(nke_close, stepwise=TRUE))

nke_weekly_fit |> pivot_longer(everything(), 
                         names_to = "Model name",
                         values_to = "Orders")
report(nke_weekly_fit)

report(nke_weekly_fit |> dplyr::select(arima110))
report(nke_weekly_fit |> dplyr::select(arima111))

glance(nke_weekly_fit) |> 
  arrange(AICc) |> 
  dplyr::select(.model:BIC)

nke_weekly_fit |>
  dplyr::select(stepwise) |>
  gg_tsresiduals()

augment(nke_weekly_fit) |>
  filter(.model=='stepwise') |>
  features(.innov, ljung_box, lag = 10, dof = 1)

nke_weekly_fit |>
  forecast(h=64) |>
  filter(.model=='stepwise') |>
  autoplot(nke_weekly_close)
```

1. Original Time Series and ACF/PACF (Plot 1):
This graph shows the weekly closing prices of Nike stock. There is a clear upward trend over time, which means the series is not stationary. The ACF plot has strong autocorrelation at many lags, especially at the beginning, and decreases slowly. The PACF plot also shows a strong spike at lag 1. These patterns suggest that the data has a strong trend and needs to be differenced to become stationary.

2. Differenced Series and ACF/PACF (Plot 2):
The second graph shows the differenced version of the series, which removes the trend. Now, the values fluctuate around zero. The ACF and PACF plots show that the autocorrelations are smaller and mostly within the blue confidence bands. This means the series is now stationary, which is a good condition for time series modeling.

3. Time Series (tsibble) and ACF/PACF (Plot 3):
This plot displays the Nike weekly closing prices using the tsibble structure. The ACF and PACF results are similar to the first plot: high autocorrelation at lag 1 and slowly decreasing values after that. These signs again confirm that the original data is non-stationary and has strong time dependence.

4. Residual Diagnostics (Plot 4):
This graph shows the residuals from the fitted ARIMA model. The residuals appear random and centered around zero over time. The ACF plot of the residuals shows that most autocorrelations are near zero and stay within the confidence limits, suggesting no clear pattern remains. The histogram of the residuals is bell-shaped, indicating they are approximately normally distributed. These are good signs of a well-fitted model.

5. Forecast Plot (Plot 5):
The last graph shows the forecast for Nike’s weekly closing prices. The dark and light blue shaded areas represent 80% and 95% confidence intervals. The forecast line is relatively flat, which means the model predicts that prices will stay stable in the short term. However, the confidence intervals become wider over time, showing that the uncertainty increases in the future.

---

# Conclusion
The analysis reveals three key findings about NKE stock:

<h4>1. Market Patterns</h4>

- Clear weekday effects (Thursday highs, Monday lows)

- January shows consistent underperformance

- Strong pre-2021 growth followed by sustained decline

<h4>2. Modeling Results</h4>

- ARIMA(1,1,1) delivers best balance of accuracy and simplicity

- All models show clean residuals (Ljung-Box p > 0.1)

- Short-term forecasts remain stable; long-term uncertain

<h4>3. Practical Applications</h4>

- Traders can exploit weekly seasonal patterns

- Investors should monitor post-2021 trend breaks

- Models perform best in stable market conditions

These results demonstrate that while NKE exhibits predictable short-term behavior, recent volatility requires cautious interpretation of long-term projections. The framework provides a foundation for ongoing analysis and decision-making.

**Note:** Some of the interpretations and explanations were refined with the help of GPT.